Constructing Sailing Match Race Schedules: Round Robin Pairing Lists

### 2015-03-25 paper.tex
Paragraph headings for the model

The first thing we model, in the first half of Figure~\ref{model:stage1}, is a
skipper. Each skipper $\sigma$ has a temporal view of his schedule in the array
of constrained integer variables $\timeSlot[0]$ to $\timeSlot[{t-1}]$, and a
state view in the array of constraint integer variables $\state[0]$ to
$\state[{f-1}]$. The state of this skipper in flight $i$ is found in
$\state[i]$, and this corresponds to timeslots $\timeSlot[{i \cdot m}]$ to
$\timeSlot[{m(i+1)-1}]$.

@@

The $\state$ variables are then used to impose match race criteria 4 (if last
in flight $i$ then not first in flight $i+1$), criteria 11 (if last in flight
$i$ then not in a bye in flight $i+1$) and criteria 12 (if flight $i$ is a bye
then flight $i$ is not first).  These criteria are imposed by \eqref{C3}, which
is a deterministic finite automata (DFA) constraint \cite{Pesant04}. The
transitions for the DFA are shown pictorially in Figure~\ref{skipper5}: arcs
represent a transition from one state to another. The DFA is encoded as a set
of transition objects $\langle q_{i},\iota,q_{j} \rangle$, where a transition
is made from state $q_{i}$ to state $q_{j}$ when encountering input $\iota$. In
addition we specify the accepting states, which are all states except the start
state (white) and the bye state (pink).  This constraint also ensures that if a
skipper has finished his matches in flight $i$ (i.e.\ $\END$) then he has
finished his matches in flight $i+1$ (and so on).

@@

A boat change occurs when a skipper has been in a $\BYE$ state in flight $i$
and then is in a match in the next flight $i+1$. This is encoded using the
array of zero/one variables \eqref{V3}, and the constraint \eqref{C4}, and
accumulated into $\totalChanges$ (\ref{V4},~\ref{C5}).

@@

We are now in a position to model criteria 1 (each skipper sails against every
other skipper) and optimisation criteria 10 (minimise boat changes).  In the
second half of Figure~\ref{model:stage1} we present a match perspective of the
schedule, using a two dimensional array of variables $\match$ \eqref{V5}, where
$\match[{i,j}]$ is the time slot that skippers $\sigma[{i}]$ and $\sigma[{j}]$
meet in a match. Only the half above the diagonal is represented, and the lower
half is made up of exactly the same variables, i.e.\ $\match[{i,j}]$ is
\emph{exactly} the same constrained variable as $\match[{j,i}]$. Constraint
\eqref{C6} states that  a match between skippers $\sigma[{i}]$ and
$\sigma[{j}]$ takes place at time $k$ (i.e.\ $\match[{i,j}] = k$) if and only
if  skipper $\sigma[{i}]$'s $k^{th}$ time slot is skipper $\sigma[{j}]$ and
conversely that skipper $\sigma[{j}]$'s $k^{th}$ time slot is skipper
$\sigma[{i}]$). Variables \eqref{V6} and constraints \eqref{C7} then convert
this from time slots to flights, i.e.\ $\modMatch[{i,j}]$ is the flight in
which skippers $\sigma[{i}]$ and $\sigma[{j}]$ meet for a match. Finally,
constraint \eqref{C8} ensures that each skipper's match occurs in different
flights. Also, since $\modMatch[{i,j}] \equiv \modMatch[{j,i}]$, setting rows
to be all different also ensures that columns are all different.

We now take a temporal view \eqref{V7}, such that $\timeVar[{\tau}]$ is a pair
$(i,j)$, stating that at time $\tau$ skippers $\sigma[{i}]$ and $\sigma[{j}]$
are in a match. We channel between the match perspective and the
temporal perspective using \eqref{C9}.

Finally we have the optimisation criteria (criteria 10) to minimise the number
of boat changes. This is handled by \eqref{V8} and \eqref{C10}.

>>>>

\paragraph{Modelling skippers:} The first thing we model, in the first half of
Figure~\ref{model:stage1}, is a skipper. Each skipper $\sigma$ has a temporal
view of his schedule in the array of constrained integer variables
$\timeSlot[0]$ to $\timeSlot[{t-1}]$, and a state view in the array of
constraint integer variables $\state[0]$ to $\state[{f-1}]$. The state of this
skipper in flight $i$ is found in $\state[i]$, and this corresponds to
timeslots $\timeSlot[{i \cdot m}]$ to $\timeSlot[{m(i+1)-1}]$.

@@

\paragraph{State transitions:} The $\state$ variables are then used to impose
match race criteria 4 (if last in flight $i$ then not first in flight $i+1$),
criteria 11 (if last in flight $i$ then not in a bye in flight $i+1$) and
criteria 12 (if flight $i$ is a bye then flight $i$ is not first).  These
criteria are imposed by \eqref{C3}, which is a deterministic finite automata
(DFA) constraint \cite{Pesant04}. The transitions for the DFA are shown
pictorially in Figure~\ref{skipper5}: arcs represent a transition from one
state to another. The DFA is encoded as a set of transition objects $\langle
q_{i},\iota,q_{j} \rangle$, where a transition is made from state $q_{i}$ to
state $q_{j}$ when encountering input $\iota$. In addition we specify the
accepting states, which are all states except the start state (white) and the
bye state (pink).  This constraint also ensures that if a skipper has finished
his matches in flight $i$ (i.e.\ $\END$) then he has finished his matches in
flight $i+1$ (and so on).

@@

\paragraph{Boat changes:} A boat change occurs when a skipper has been in a
$\BYE$ state in flight $i$ and then is in a match in the next flight $i+1$.
This is encoded using the array of zero/one variables \eqref{V3}, and the
constraint \eqref{C4}, and accumulated into $\totalChanges$
(\ref{V4},~\ref{C5}).

@@

\paragraph{Match perspective:} We are now in a position to model criteria 1
(each skipper sails against every other skipper) and optimisation criteria 10
(minimise boat changes).  In the second half of Figure~\ref{model:stage1} we
present a match perspective of the schedule, using a two dimensional array of
variables $\match$ \eqref{V5}, where $\match[{i,j}]$ is the time slot that
skippers $\sigma[{i}]$ and $\sigma[{j}]$ meet in a match. Only the half above
the diagonal is represented, and the lower half is made up of exactly the same
variables, i.e.\ $\match[{i,j}]$ is \emph{exactly} the same constrained
variable as $\match[{j,i}]$. Constraint \eqref{C6} states that  a match between
skippers $\sigma[{i}]$ and $\sigma[{j}]$ takes place at time $k$ (i.e.\
$\match[{i,j}] = k$) if and only if  skipper $\sigma[{i}]$'s $k^{th}$ time slot
is skipper $\sigma[{j}]$ and conversely that skipper $\sigma[{j}]$'s $k^{th}$
time slot is skipper $\sigma[{i}]$). Variables \eqref{V6} and constraints
\eqref{C7} then convert this from time slots to flights, i.e.\
$\modMatch[{i,j}]$ is the flight in which skippers $\sigma[{i}]$ and
$\sigma[{j}]$ meet for a match. Finally, constraint \eqref{C8} ensures that
each skipper's match occurs in different flights. Also, since $\modMatch[{i,j}]
\equiv \modMatch[{j,i}]$, setting rows to be all different also ensures that
columns are all different.

\paragraph{Temporal perspective:} We now take a temporal view \eqref{V7}, such
that $\timeVar[{\tau}]$ is a pair $(i,j)$, stating that at time $\tau$ skippers
$\sigma[{i}]$ and $\sigma[{j}]$ are in a match. We channel between the match
perspective and the temporal perspective using \eqref{C9}.

\paragraph{Optimisation criteria:} Finally we have the optimisation criteria
(criteria 10) to minimise the number of boat changes. This is handled by
\eqref{V8} and \eqref{C10}.
